Verification of Stencil Computations through Spatial Specifications

### 2016-07-07 paper.tex
edits

\documentclass[9pt]{sigplanconf}

@@

\item we evaluate our approach with an
implementation as an extension of CamFort,
an open-source program analysis tool for Fortran
programs. We use our implementation to evaluate our approach against
a corpus of scientific applications, totalling roughly one million
lines of code;

@@

\subsection{Notation and convention}
\label{sec:notation}

\renewcommand*{\arraystretch}{0.8}
For the target language
syntax, $e$ ranges over its expressions (may be impure) and $v$
overs its (imperative) variables. %By
%``variable'', we mean in the imperative sense (named binders to memory cells).

@@

\begin{figure}[t]
\vspace{-0.9em}
\input{region-model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Model of regions,
$\interp{-}_n : \textit{region} \rightarrow \mathcal{P}(\domainVal{}^n)$}
\label{fig:region-model}
\vspace{-0.7em}
\end{figure}

@@

\begin{figure}
\vspace{-0.5em}
\input{spatial-model.tex}
\caption{\textit{spec} model,
$\interp{-}_n : \textit{spec} \rightarrow \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$}
\label{fig:spatial-model}
\vspace{-0.75em}
\end{figure}

@@

\noindent
We briefly sketch the algorithms for inferring and synthesising
specifications (Sect.~\ref{subsec:inference}), for checking code
against specifications (Sect.~\ref{subsec:checking}) and for
synthesising programs from specifications
(Sect.~\ref{subsec:synthesis}).  These utilise the semantic model of
Section~\ref{sec:semantics}.

@@

The inference relies on some standard program analyses, computed
before the main inference procedure:
%
\begin{enumerate}
\item basic blocks (CFG);
\item induction variables per basic block;
\item (interprocedural) data-flow analysis, providing a \emph{flows to}
  graph (as shorthand, the function
  $\mathsf{flowTo}$ is used, implicitly parameterised by this graph,
  mapping an expression to the set of all expressions
  with forwards data-flow to this expression, based on the transitive
  closure of the flows graph);
\item type information per variable, where we use the predicate
\arrayTy{} to classify variables of array type.
\end{enumerate}
%

@@

\begin{figure}[t]
\vspace{-0.6em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
specification inference}
\label{fig:inference-steps-informal}
\vspace{-0.8em}
\end{figure}

\noindent
\textit{1).\;} Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\begin{equation*}
P(M) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow M\}
\end{equation*}
%
This corresponds to all rotations of the space.  For our example:
%
\begin{align*}
P(M_0) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation flips the order of the two elements in the 
vectors. \\

@@

%\newcommand{\mkBig}[2]{\overset{\LARGE{\term{#1}}}{#2}}
\begin{definition} The $\odot$ operation on $\textsf{A}(\specDNF{})$
  is defined in terms of the intermediate operator $\hat{\odot}$ on
  \specDNF{}, defined:
%%
\begin{align*}
S \,\hat{\odot}\, \epsilon = S \quad
\epsilon \,\hat{\odot}\, S = S \qquad
S \,\hat{\odot}\, T = \bplus_{(\tyProd{s}, \tyProd{t}) \in (S \times
   T)} (\tyProd{s} \; \term{*} \; \tyProd{t})
\end{align*}
%%
That is, the $\hat{\odot}$ product of two specifications $S$ and $T$ in DNF form 
is the $\term{+}$ sum of all pairwise $\term{*}$ products for every
pair of $\tyProd{spec}$ drawn from $S$ and $T$. The result
is indeed in DNF-form, where $\hat{\odot}$ is essentially applying
 the (\trule{\textsc{dist}}) rule of our equational theory
 (Fig.~\ref{fig:equations},p.~\pageref{fig:equations}). The operation is
commutative and associative, and sound with respect to our
model (see Lemma~\ref{lem:alg-soundness} below).
The $\hat{\odot}$ operation is then lifted to all combinations of
\textsf{Approx} for $\odot$:
\begin{align*}
%\textsf{lower}(S) \odot \textsf{lower}(T) = \; & 
%      \textsf{lower} \, (S \hat{\odot} T) \\
%\textsf{upper}(S) \odot \textsf{upper}(T) = \; &
%      \textsf{upper} \, (S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{lower}(S) = \;
       \textsf{both} \, (S \hat{\odot} T, \; T) \\
\textsf{upper}(S) \odot \textsf{exact}(T) = \; & 
       %\textsf{exact}(T) \odot \textsf{upper}(S) =
      \; \textsf{both} \, (T, \; S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{both}(T_l,T_u) = \; &
        %\textsf{both}(T_l,T_u) \odot \textsf{lower}(S) = \;
        \textsf{both} \, (S \hat{\odot} T_l, \; T_u) \\
\textsf{upper}(S) \odot \textsf{both}(T_l,T_u) = \; & \textsf{both} \,
                                                      (T_l, \; S \hat{\odot} T_u) \\
\textsf{both}(S_l,S_u) \odot
\textsf{both}(T_l,T_u) & = \textsf{both} (S_l \hat{\odot} T_l, \; S_u
                         \hat{\odot} T_u) \\
\textit{inj}(S) \odot \textit{inj}(T) = \; & \textit{inj}(S
                                              \hat{\odot} T)
\end{align*}
where in the last case $\textit{inj}$ corresponds to the unary
injections of \textsf{Approx}. For brevity we omit the cases where the parameters above are flipped
as $\odot$ is commutative.
\end{definition}

@@

Package   & {Total} & {Parsed} & {Total} & {Parsed}  \\ % braces prevent headers from being formatted as numbers
\hline
UM        & \umFiles                       & \umparseOk                      & \umLoC                      & \umlinesParsed \\
E3MG      & \ethreemgeaFiles               & \ethreemgeaparseOk              & \ethreemgeaLoC              & \ethreemgealinesParsed \\
BLAS      & \blasFiles                     & \blasparseOk                    & \blasLoC                    & \blaslinesParsed \\
Hybrid4   & \hybridfourFiles               & \hybridfourparseOk              & \hybridfourLoC              & \hybridfourlinesParsed \\
GEOS-Chem & \geoschemFiles                 & \geoschemparseOk                & \geoschemLoC                & \geoschemlinesParsed \\
Navier    & \navierFiles                   & \navierparseOk                  & \navierLoC                  & \navierlinesParsed \\
CP        & \computationalphysicstwoFiles  & \computationalphysicstwoparseOk & \computationalphysicstwoLoC & \computationalphysicstwolinesParsed \\
\hline
\textbf{Total} & \overallFiles  & \overallparseOk & \overallLoC & \overalllinesParsed \\
\hline
\end{tabular}
\caption{Summary of software packages used for
  evaluation\label{tab:corpus}}
\vspace{-0.7em}
\end{table}

@@

\begin{table}
\sisetup{table-format=4,table-text-alignment=right}
\footnotesize
\centering
\begin{tabular}{lS@{~~~}rS@{~~~}rSS[table-format=1]}
\hline
  & \multicolumn{2}{c}{Potential} & \multicolumn{2}{c}{Actual} & {Exact} & {Inexact}    \\
{Package}     & {\#} & \si{\percent} & {\#} & \si{\percent} \\
\hline
UM            & \umtickAssign & \SI{\umtickAssignPercent}{\percent} &
                \umtickAssignSuccess & \SI{\umtickAssignSuccessPercent}{\percent} &
                \umnumStencilLines &
                \umboundedBoth  \\
E3MG          & \ethreemgeatickAssign & \SI{\ethreemgeatickAssignPercent}{\percent} &
                \ethreemgeatickAssignSuccess & \SI{\ethreemgeatickAssignSuccessPercent}{\percent} &
                \ethreemgeanumStencilLines &
                \ethreemgeaboundedBoth \\
BLAS          & \blastickAssign & \SI{\blastickAssignPercent}{\percent} &
                \blastickAssignSuccess & \SI{\blastickAssignSuccessPercent}{\percent} &
                \blasnumStencilLines &
                \blasboundedBoth \\
Hybrid4       & \hybridfourtickAssign & \SI{\hybridfourtickAssignPercent}{\percent} &
                \hybridfourtickAssignSuccess & \SI{\hybridfourtickAssignSuccessPercent}{\percent} &
                \hybridfournumStencilLines &
                \hybridfourboundedBoth\\
GEOS-Chem     & \geoschemtickAssign & \SI{\geoschemtickAssignPercent}{\percent} &
                \geoschemtickAssignSuccess & \SI{\geoschemtickAssignSuccessPercent}{\percent} &
                \geoschemnumStencilLines &
                \geoschemboundedBoth \\
Navier        & \naviertickAssign & \SI{\naviertickAssignPercent}{\percent} &
                \naviertickAssignSuccess & \SI{\naviertickAssignPercent}{\percent} &
                \naviernumStencilLines &
                \navierboundedBoth \\
CP            & \computationalphysicstwotickAssign & \SI{\computationalphysicstwotickAssignPercent}{\percent} &
                \computationalphysicstwotickAssignSuccess & \SI{\computationalphysicstwotickAssignSuccessPercent}{\percent} &
                \computationalphysicstwonumStencilLines &
                \computationalphysicstwoboundedBoth \\
\hline
\textbf{Total}& \overalltickAssign & \SI{\overalltickAssignPercent}{\percent}  &
                \overalltickAssignSuccess & \SI{\overalltickAssignSuccessPercent}{\percent} &
                \overallnumStencilLines &
                \overallboundedBoth    \\
\hline
\end{tabular}
\caption{Stencil specification assignment rates\label{tab:tickAssign}}
\vspace{-0.8em}
\end{table}

@@

The Unified Model has an implementation of the Smagorinsky
subgrid-scale model for calculating turbulence on which our
inference yields 39 specs from 340 lines of code. One example is:

\begin{minted}[breakindent=0em,linenos=false,xleftmargin=0em,breakafter=)]{fortran}
!=stencil readOnce,(forward(depth=1,dim=3,irreflexive))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(reflexive(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2,irreflexive))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(reflexive(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,irreflexive))*(reflexive(dim=3))+(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2))*(reflexive(dim=3))::x
\end{minted}
%
This specifies the access pattern to a 3-dimensional array (which we have renamed to be called \texttt{x}). The array is accessed 8 times in the computation. We tried introducing indexing errors by changing the offset in one dimension of one of the references to \texttt{x}. For example changing \mintinline{fortran}{x(i+1,j,k)} to \mintinline{fortran}{x(i,j,k)} or \mintinline{fortran}{x(i-1,j,k)}.

There are 48 possible errors of this type of which we found only 6 were not 
detected by the specification checking. The reason is that the inferred
specification describes a family of access patterns. Some offset errors
produce a stencil computation that belongs to the same family and goes
undetected.
One of the reasons this specification can catch many offset errors is that it is
a \texttt{readOnce} specification. This prevents offset errors leading to
duplicate access patterns be caught even if the resulting set of accesses are
part of the same family.
% /um/trunk/src/atmosphere/diffusion_and_filtering/turb_smagorinsky.F90
% Tried introducing indexing errors replacing one of the indices of cx_rho_instance with -1,0,+1
% 42 cases were caught
% 6 cases were not caught
\section{Related work}
\label{sec:discussion}

>>>>

\documentclass[9pt,preprint]{sigplanconf}

@@

\item we provide an implementation of our approach as an extension to
  CamFort, an open-source program analysis tool for Fortran programs.
  We use this for a quantiative evaluation of our approach over a corpus of
  scientific applications, totalling roughly one million lines of
  code;

@@

\paragraph{Notation and convention}
%\label{sec:notation}

\renewcommand*{\arraystretch}{0.8}
For the target language, $e$ ranges over expressions (may be impure) and $v$
overs its (imperative) variables. %By
%``variable'', we mean in the imperative sense (named binders to memory cells).

@@

\begin{figure}[t]
\vspace{-0.9em}
\input{region-model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Model of regions,
$\interp{-}_n : \textit{region} \rightarrow \mathcal{P}(\domainVal{}^n)$}
\label{fig:region-model}
\vspace{-1em}
\end{figure}

@@

\begin{figure}
\input{spatial-model.tex}
\caption{\textit{spec} model,
$\interp{-}_n : \textit{spec} \rightarrow \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$}
\label{fig:spatial-model}
\vspace{-0.75em}
\end{figure}

@@

\noindent
We briefly sketch the algorithms for inferring and synthesising
specifications (Sect.~\ref{subsec:inference}), for checking code
against specifications (Sect.~\ref{subsec:checking}). These utilise the semantic model of
Section~\ref{sec:semantics}.

@@

The inference relies on some standard program analyses, computed
before the main inference procedure:
%
\textbf{(1)} basic blocks (CFG);
\textbf{(2)} induction variables per basic block;
\textbf{(3)} (interprocedural) data-flow analysis, providing a \emph{flows to}
  graph (as shorthand, the function
  $\mathsf{flowTo}$ is used, implicitly parameterised by this graph,
  mapping an expression to the set of all expressions
  with forwards data-flow to this expression, based on the transitive
  closure of the flows graph);
\textbf{(4)} type information per variable, where we use the predicate
\arrayTy{} to classify variables of array type.
%

@@

\begin{figure}[t]
\vspace{-0.8em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
specification inference}
\label{fig:inference-steps-informal}
\vspace{-0.8em}
\end{figure}

\noindent
\textit{1).\;} Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\begin{equation*}
P(M) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow M\}
\end{equation*}
%
This corresponds to all rotations of the space.  For our example:
%
\begin{align*}
P(M_0) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation flips the order of the two elements in the 
vectors. \\[-1em]

@@

%\newcommand{\mkBig}[2]{\overset{\LARGE{\term{#1}}}{#2}}
\begin{definition} The $\odot$ operation on $\textsf{A}(\specDNF{})$
  is defined in terms of the intermediate operator $\hat{\odot}$ on
  \specDNF{}, defined:
%%
\begin{align*}
S \,\hat{\odot}\, \epsilon = S \quad
\epsilon \,\hat{\odot}\, S = S \qquad
S \,\hat{\odot}\, T = \bplus_{(\tyProd{s}, \tyProd{t}) \in (S \times
   T)} (\tyProd{s} \; \term{*} \; \tyProd{t})
\end{align*}
%%
That is, the $\hat{\odot}$ product of two specifications $S$ and $T$ in DNF form 
is the $\term{+}$ sum of all pairwise $\term{*}$ products for every
pair of $\tyProd{spec}$ drawn from $S$ and $T$. The result
is indeed in DNF-form, where $\hat{\odot}$ is essentially applying
 the (\trule{\textsc{dist}}) rule of our equational theory
 (Fig.~\ref{fig:equations},p.~\pageref{fig:equations}). The operation is
commutative and associative, and sound with respect to our
model (see Lemma~\ref{lem:alg-soundness} below).
The $\hat{\odot}$ operation is then lifted to all combinations of
\textsf{Approx} for $\odot$:
\begin{align*}
\textsf{lower}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{lower}(S) = \;
       \textsf{both} \, (S \hat{\odot} T, \; T) \\
\textsf{upper}(S) \odot \textsf{exact}(T) = \; & 
       %\textsf{exact}(T) \odot \textsf{upper}(S) =
       \textsf{both} \, (T, \; S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{both}(T_l,T_u) = \; &
        %\textsf{both}(T_l,T_u) \odot \textsf{lower}(S) = \;
        \textsf{both} \, (S \hat{\odot} T_l, \; T_u) \\
\textsf{upper}(S) \odot \textsf{both}(T_l,T_u) = \; & \textsf{both} \,
                                                      (T_l, \; S \hat{\odot} T_u) \\
\textsf{both}(S_l,S_u) \odot
\textsf{both}(T_l,T_u) & = \textsf{both} (S_l \hat{\odot} T_l, \; S_u
                         \hat{\odot} T_u) \\
\textit{inj}(S) \odot \textit{inj}(T) = \; & \textit{inj}(S
                                              \hat{\odot} T)
\end{align*}
where in the last case $\textit{inj}$ corresponds to the unary
injections of \textsf{Approx}. For brevity we omit the cases where the parameters above are flipped
as $\odot$ is commutative.
\end{definition}

@@

Package   & {Total} & {Parsed} & {Total} & {Parsed}  \\ % braces prevent headers from being formatted as numbers
\hline
UM        & \umFiles                       & \umparseOk                      & \umLoC                      & \umlinesParsed \\
E3MG      & \ethreemgeaFiles               & \ethreemgeaparseOk              & \ethreemgeaLoC              & \ethreemgealinesParsed \\
BLAS      & \blasFiles                     & \blasparseOk                    & \blasLoC                    & \blaslinesParsed \\
Hybrid4   & \hybridfourFiles               & \hybridfourparseOk              & \hybridfourLoC              & \hybridfourlinesParsed \\
GEOS-Chem & \geoschemFiles                 & \geoschemparseOk                & \geoschemLoC                & \geoschemlinesParsed \\
Navier    & \navierFiles                   & \navierparseOk                  & \navierLoC                  & \navierlinesParsed \\
CP        & \computationalphysicstwoFiles  & \computationalphysicstwoparseOk & \computationalphysicstwoLoC & \computationalphysicstwolinesParsed \\
\hline
\textbf{Total} & \overallFiles  & \overallparseOk & \overallLoC & \overalllinesParsed \\
\end{tabular}
\caption{Summary of software packages used for
  evaluation\label{tab:corpus}}
\vspace{-0.7em}
\end{table}

@@

\begin{table}
\sisetup{table-format=4,table-text-alignment=right}
\footnotesize
\centering
\begin{tabular}{lS@{~~~}rS@{~~~}rSS[table-format=1]}
\hline
  & \multicolumn{2}{c}{Potential} & \multicolumn{2}{c}{Actual} & {Exact} & {Inexact}    \\
{Package}     & {\#} & \si{\percent} & {\#} & \si{\percent} \\
\hline
UM            & \umtickAssign & \SI{\umtickAssignPercent}{\percent} &
                \umtickAssignSuccess & \SI{\umtickAssignSuccessPercent}{\percent} &
                \umnumStencilLines &
                \umboundedBoth  \\
E3MG          & \ethreemgeatickAssign & \SI{\ethreemgeatickAssignPercent}{\percent} &
                \ethreemgeatickAssignSuccess & \SI{\ethreemgeatickAssignSuccessPercent}{\percent} &
                \ethreemgeanumStencilLines &
                \ethreemgeaboundedBoth \\
BLAS          & \blastickAssign & \SI{\blastickAssignPercent}{\percent} &
                \blastickAssignSuccess & \SI{\blastickAssignSuccessPercent}{\percent} &
                \blasnumStencilLines &
                \blasboundedBoth \\
Hybrid4       & \hybridfourtickAssign & \SI{\hybridfourtickAssignPercent}{\percent} &
                \hybridfourtickAssignSuccess & \SI{\hybridfourtickAssignSuccessPercent}{\percent} &
                \hybridfournumStencilLines &
                \hybridfourboundedBoth\\
GEOS-Chem     & \geoschemtickAssign & \SI{\geoschemtickAssignPercent}{\percent} &
                \geoschemtickAssignSuccess & \SI{\geoschemtickAssignSuccessPercent}{\percent} &
                \geoschemnumStencilLines &
                \geoschemboundedBoth \\
Navier        & \naviertickAssign & \SI{\naviertickAssignPercent}{\percent} &
                \naviertickAssignSuccess & \SI{\naviertickAssignPercent}{\percent} &
                \naviernumStencilLines &
                \navierboundedBoth \\
CP            & \computationalphysicstwotickAssign & \SI{\computationalphysicstwotickAssignPercent}{\percent} &
                \computationalphysicstwotickAssignSuccess & \SI{\computationalphysicstwotickAssignSuccessPercent}{\percent} &
                \computationalphysicstwonumStencilLines &
                \computationalphysicstwoboundedBoth \\
\hline
\textbf{Total}& \overalltickAssign & \SI{\overalltickAssignPercent}{\percent}  &
                \overalltickAssignSuccess & \SI{\overalltickAssignSuccessPercent}{\percent} &
                \overallnumStencilLines &
                \overallboundedBoth   
\end{tabular}
\caption{Stencil specification assignment rates\label{tab:tickAssign}}
\vspace{-1em}
\end{table}

@@

\begin{figure}[t]
\begin{minted}[breakindent=0em,linenos=false,xleftmargin=0em,breakafter=)]{fortran}
!=stencil readOnce,(forward(depth=1,dim=3,irreflexive))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(reflexive(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2,irreflexive))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(reflexive(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,irreflexive))*(reflexive(dim=3))+(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2))*(reflexive(dim=3))::x
\end{minted}
\caption{A complex specification inferred from
  \textbf{Unified Model}\label{fig:smagorinsky}}
\vspace{-1em}
\end{figure}

\noindent
The Unified Model has an implementation of the Smagorinsky
subgrid-scale model for calculating turbulence on which our
inference yields 39 specs from 340 lines of code. We show
one example in Figure~\ref{fig:smagorinsky}. 
This specifies the access pattern to a 3-dimensional array (which we have renamed to be called \texttt{x}). The array is accessed 8 times in the computation. We tried introducing indexing errors by changing the offset in one dimension of one of the references to \texttt{x}. For example changing \mintinline{fortran}{x(i+1,j,k)} to \mintinline{fortran}{x(i,j,k)} or \mintinline{fortran}{x(i-1,j,k)}.

There are 48 possible errors of this type of which only 6 were not 
detected by the checker. The reason is that the inferred
specification describes a family of access patterns. Some offset errors
produce a stencil computation that belongs to the same family and goes
undetected.
One of the reasons this specification can catch many offset errors is that it is
a \texttt{readOnce} specification. This prevents offset errors leading to
duplicate access patterns be caught even if the resulting set of accesses are
part of the same family.
% /um/trunk/src/atmosphere/diffusion_and_filtering/turb_smagorinsky.F90
% Tried introducing indexing errors replacing one of the indices of cx_rho_instance with -1,0,+1
% 42 cases were caught
% 6 cases were not caught
\section{Related work}
\label{sec:discussion}
